---
title: "Plotly Interactive Graphics"
author: "Connor Guest"
date: "6/1/2019"
output: 
  html_document:
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, include = FALSE}
# install.packages("plotly")
# install.packages("tidyverse")

library(plotly)
library(readr)
library(dplyr)
library(tidyverse)
library(readxl)
```

```{r, include = FALSE}
msci <- read_csv("MSCI.csv")

names(msci)
head(msci)
```

### Learning Plotly, a library for interactive graphics

#### Here is my first graphic

In this graph we are plotting the MSCI data taken from here: https://finance.yahoo.com/quote/MSCI/history?period1=1527807600&period2=1559343600&interval=1d&filter=history&frequency=1d

```{r plotly line chart}
# Create a times series plot of Open against Date
msci %>% 
    plot_ly(x = ~Date, y = ~Open) %>% 
    add_lines()
```

```{r}
# Create a scatterplot with Date on the x-axis and Volume on the y-axis
msci %>% 
  plot_ly(x = ~Date, y = ~Volume) %>%
  add_markers()
```

Happiness Report

```{r, include = FALSE}
happy <- read_excel("HappinessData.xls")
countries <- read_csv("countries of the world.csv")

countries <- rename(countries, country = "Country", population = "Population", region = "Region", gdp = "GDP ($ per capita)" )
countries <- select(countries, country, population, region, gdp)
countries <- countries %>%
              mutate(
                income = case_when(
                  gdp > 16000 ~ "high",
                  gdp > 9000 ~ "upper-middle",
                  gdp > 2000 ~ "lower-middle",
                  TRUE        ~ "low"
                )
              )


summary(happy)
str(happy)
glimpse(happy)
names(happy)

glimpse(countries)

# join happy data set with countries 

happycountries <- happy %>% 
                  left_join(countries, by = c("Country name" = "country"))


happycountries <- rename(happycountries, country = "Country name", happiness = "Life Ladder", life.expectancy = "Healthy life expectancy at birth", log.gdp = "Log GDP per capita", gdp.per.capita = "gdp")

```

```{r}
# Create a coded scatterplot of happiness vs. life.expectancy for 2018

happycountries %>%
  filter(Year == 2018) %>% 
  plot_ly(x = ~life.expectancy, y = ~happiness) %>%
  add_markers(color = ~region, size = ~log.gdp)

```

```{r}
# Plotting symbols and polishing the graphic with tooltip and labels
happycountries %>%
  filter(Year == 2018) %>%
  plot_ly(x = ~life.expectancy, y = ~happiness,
  hoverinfo = "text",
  text = ~paste("Country: ", country,
  "<br> Income: ", income,
  "<br> Happiness: ", round(happiness, 2),
  "<br> Life Expectancy: ", round(life.expectancy, 2),
  "<br> GDP Per Capita: ", round(gdp.per.capita, 2))) %>%
  add_markers(symbol = ~income, 
              symbols = c("circle-open", "square-open", "star-open", "x-thin-open")) %>%
  layout(xaxis = list(title = "Life Expectancy"),
         yaxis = list(title = "National Happiness Score"))

```

Import new data set
```{r, include=FALSE}
us_economy <- read_csv("state_economic_data.csv")
```

```{r}

# Create a scatter plot of house_price over time by state
us_economy %>%
  filter(year == 2017) %>%
  plot_ly(x = ~gdp, y = ~house_price) %>%
  add_markers(size = ~population, 
              color = ~region, 
              marker = list(sizemode = "diameter"))

# Create a line chart of house_price over time by state
us_economy %>%
  filter(year >= 2000) %>%
  group_by(state) %>%
  plot_ly(x = ~year, y = ~house_price) %>%
  add_lines()
```
Adding Animation

```{r}
# Create an animated bubble chart of house_price against gdp
us_economy %>%
  plot_ly(
    x = ~gdp, y = ~house_price,
    hoverinfo = "text", text = ~state
  ) %>%
  add_markers(
    size = ~population, color = ~region, 
    frame = ~year, ids = ~state,
    marker = list(sizemode = "diameter", sizeref = 2) # size ref changes the size of the bubble
  ) %>%
  layout(
    xaxis = list(title = "Real GDP (millions USD)", type = "log"),
    yaxis = list(title = "Housing price index")
  )
```
```{r, include=FALSE}
# Factors as frames
# Time is not the only type of variable that can be used to specify the frames of an animation. You can also create an animation that cycles through subplots created by factors rather than using a grid layout.
# 
# In this exercise, you will adapt your answer from the previous question to create an animation that cycles through bubble charts for each region in 2017.
# 
# Note: The default settings for the animation make it go very quickly. In the next lesson, you'll learn how to adjust the speed. For now, you can use the slider bar to navigate between regions if you want to investigate the relationships.
# 
# plotly and the us_economy data set have already been loaded.
```



```{r}
# Animate a bubble chart of house_price against gdp over region
ani <- us_economy %>%
  filter(year == 2017) %>%
  plot_ly(x = ~gdp, y = ~house_price) %>%
  add_markers(size = ~population, color = ~region, 
              frame = ~region, ids = ~state,
              marker = list(sizemode = "diameter"))
```

Animation Attributes 
https://github.com/plotly/plotly.js/blob/master/src/plots/animation_attributes.js
```{r}
# ani is the plot created above
ani %>% 
   animation_opts(
     frame = 2000, 
     transition = 300, #this is how long the frame takes
     easing = "elastic"
   ) %>%
   animation_slider( 
     currentvalue = list(
       prefix = NULL, # this get rid of the slider 
       font = list(color = "red") # changes the color on the slider
     )
   ) %>%   
   layout(
     xaxis = list(title = "Real GDP (millions USD)"),
     yaxis = list(title = "Housing price index")
   )
```

Adding background text

```{r}
# Add the year as background text and remove the slider
us_economy %>%
  plot_ly(x = ~gdp, y = ~house_price, hoverinfo = "text", text = ~state) %>%
  add_text(x = 200000, y = 450, text = ~year, frame = ~year,                  # this is where the code is
           textfont = list(color = toRGB("gray80"), size = 150)) %>%
  add_markers(size = ~population, color = ~region, 
              frame = ~year, ids = ~state,
              marker = list(sizemode = "diameter", sizeref = 3)) %>%
  layout(xaxis = list(title = "Real GDP (millions USD)", type = "log"),
         yaxis = list(title = "Housing price index")) %>%
  animation_slider(hide = TRUE)
```
Layering
Plotting the baseline - keeping dots in the background
You can use different data sets for different layers in plotly, just like you can in ggplot2. This is a useful strategy when you want to display a baseline time point (or group) in an animation. Here is an animated scatterplot of housing price index against real GDP, keeping the 1997 data points in the background.

```{r}
# extract the 1997 data
us1997 <- us_economy %>%
  filter(year == 1997)

# create an animated scatterplot with baseline from 1997
us_economy %>%
  plot_ly(x = ~gdp, y = ~house_price) %>%
  add_markers(data = us1997, marker = list(color = toRGB("gray60"), opacity = 0.5)) %>%
  add_markers(frame = ~year, ids = ~state, data = us_economy, showlegend = FALSE, alpha = 0.5) %>%
  layout(xaxis = list(type = "log"))
```



